### 2023-10-17 docs/submission/documents_august_submission/diff/template_Article.tex

\documentclass[]{article}

%opening
\title{}
\author{}

\begin{document}

\maketitle

Given the large number of changes to the manuscript recommended by the reviewers, including a major overhaul of the paper structure, visual comparison packages such as \texttt{latexdiff} cannot generate a meaningful or informative marked-up version of the revised manuscript. We hope that the author response to the referee comments, together with the revised manuscript, are together sufficient to describe the major changes to the manuscript in this revision.

\end{document}